---
title: Quality control on the CHART-seq data
author: Aaron Lun
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    fig_caption: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(BiocStyle)
knitr::opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
```

# Setting up the data

Defining the BAM files to use here:

```{r}
bam.files <- list.files("../bam", full=TRUE, pattern="\\.bam$")
prefixes <- sub("\\.bam$", "", basename(bam.files))
prefixes
```

# Getting read alignment statistics

Reporting alignment statistics for each file:

```{r}
library(Rsamtools)
diagnostics <- list()
for (idx in seq_along(bam.files)) {
	bam <- bam.files[idx]
    everything <- scanBam(bam.files[idx], param=ScanBamParam(
        what=c("mapq", "flag")))[[1]]

    total <- length(everything$flag)
    mapped <- bitwAnd(everything$flag, 0x4)==0
    marked <- bitwAnd(everything$flag, 0x400)!=0
    filtered <- everything$mapq >= 10
    survivor <- filtered & !marked
                
    diagnostics[[prefixes[idx]]] <- c(Total=total, Mapped=sum(mapped), 
        Filtered=sum(filtered, na.rm=TRUE), Marked=sum(marked), 
        Remaining=sum(survivor, na.rm=TRUE))
}
diag.stats <- data.frame(do.call(rbind, diagnostics))
diag.stats$Prop.mapped <- diag.stats$Mapped/diag.stats$Total*100
diag.stats$Prop.marked <- diag.stats$Marked/diag.stats$Mapped*100
diag.stats$Prop.left <- diag.stats$Filtered/diag.stats$Total*100
diag.stats
```

# Getting paired read statistics

Getting the insert sizes for each BAM file:

```{r}
library(csaw)
param <- readParam(dedup=TRUE, minq=10, pe="both")

pe.diagnostics <- list()
par(mfrow=c(1,2))    
for (idx in seq_along(bam.files)) {
    out <- getPESizes(bam.files[idx], param=param)
    toshow <- out$sizes
    toshow <- toshow[toshow <= 1e3]
    hist(toshow, breaks=100, xlab="Insert size (bp)",
        main=prefixes[idx])
    pe.diagnostics[[prefixes[idx]]] <- out$diagnostics
}
```

Reporting the overall PE diagnostics:

```{r}
do.call(rbind, pe.diagnostics)
```

# Getting cross-correlations

Computing cross-correlations between forward and reverse strands.
This requires that we only use one read from each pair, otherwise it would be trivial due to the read pair insert size.

```{r}
param <- readParam(dedup=TRUE, minq=10, pe="first")
par(mfrow=c(1,2))    
for (idx in seq_along(bam.files)) {
    out <- correlateReads(bam.files[idx], param=param)
    plot(seq_along(out)-1, out, xlab="Delay (bp)", ylab="CCF",
        main=prefixes[idx])            
}
```

# Session information

```{r}
sessionInfo()
```
