---
title: Consolidating DB analyses for H3K4me3 cut-and-run 
author: Aaron Lun
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    fig_caption: yes
---

# Reading in the results

Here we consolidate the various window-based analyses into a single set of results.
First, we define the files to load in:

```{r}
library(csaw)
all.files <- list.files(full=TRUE, pattern="^h3k4me3_.*rds$")
all.files
```

And we load in the results:

```{r}
all.results <- lapply(all.files, readRDS)
all.regions <- lapply(all.results, "[[", i="windows")
all.rnai <- lapply(all.results, "[[", i="RNAi")
all.lna <- lapply(all.results, "[[", i="LNA")
```

We apply an IUT on the one-sided p-values.
This involves converting each p-value into two one-sided values, performing IUTs on each side, and then effectively performing a Bonferoni correction to combine them again.

```{r}
all.combined <- vector("list", length(all.results))
for (idx in seq_along(all.rnai)) {
    cur.rnai <- all.rnai[[idx]]
    cur.lna <- all.lna[[idx]]

    up.rnai <- cur.rnai$logFC > 0
    left.rnai <- ifelse(up.rnai, cur.rnai$PValue/2, 1-cur.rnai$PValue/2)
    right.rnai <- ifelse(!up.rnai, cur.rnai$PValue/2, 1-cur.rnai$PValue/2)

    up.lna <- cur.lna$logFC > 0
    left.lna <- ifelse(up.lna, cur.lna$PValue/2, 1-cur.lna$PValue/2)
    right.lna <- ifelse(!up.lna, cur.lna$PValue/2, 1-cur.lna$PValue/2)

    left.p <- pmax(left.rnai, left.lna)
    right.p <- pmax(right.rnai, right.lna)
    all.combined[[idx]] <- data.frame(logFC=(cur.rnai$logFC + cur.lna$logFC)/2,
        PValue=pmin(left.p, right.p, 0.5)*2)
}
```

# Reporting DB regions

## RNAi

We consolidate all of these results into a single set of DB regions, with a maximum region width of 10 kbp.
First we do so for RNAi:

```{r}
out <- consolidateSizes(all.regions, all.rnai, 
    merge.args=list(max.width=10e3, tol=100))
out$region
```

We examine the number of regions that change, and in which direction.

```{r}
table(out$table$FDR <= 0.05, out$table$direction)
```

We also get the best log-fold changes in each region.

```{r}
best <- getBestTest(unlist(out$id), do.call(rbind, all.rnai))
head(best)
```

Adding annotation.

```{r}
library(org.Hs.eg.db)
library(EnsDb.Hsapiens.v86)
fiddling <- out$region
seqlevels(fiddling) <- sub("^chr", "", seqlevels(fiddling))
suppressWarnings({
    surrounds <- detailRanges(orgdb=org.Hs.eg.db, txdb=EnsDb.Hsapiens.v86, 
        fiddling, key.field="ENSEMBL")
})
```

We save this result to file.

```{r}
output <- data.frame(as.data.frame(out$region)[,1:3],
    out$table, Best.logFC=best$logFC, surrounds)
output <- output[order(output$PValue),]    
write.table(file="RNAi_h3k4me3_regions.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
head(output)
```

Also the significant hits to a BED file for viewing.

```{r}
db.regions <- out$region
db.regions$score <- -10*log10(out$table$PValue)
library(rtracklayer)
export(con='RNAi_h3k4me3.bed', db.regions[out$table$FDR<=0.05])
```

## LNA

We repeat this process for LNA:

```{r}
out <- consolidateSizes(all.regions, all.lna, 
    merge.args=list(max.width=10e3, tol=100))
```

We examine the number of regions that change, and in which direction.

```{r}
table(out$table$FDR <= 0.05, out$table$direction)
```

We also get the best log-fold changes in each region.

```{r}
best <- getBestTest(unlist(out$id), do.call(rbind, all.lna))
head(best)
```

We save this result to file (no need to recompute annotation, as the regions are the same).

```{r}
output <- data.frame(as.data.frame(out$region)[,1:3],
    out$table, Best.logFC=best$logFC, surrounds)
output <- output[order(output$PValue),]    
write.table(file="LNA_h3k4me3_regions.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
```

Also the significant hits to a BED file for viewing.

```{r}
db.regions <- out$region
db.regions$score <- -10*log10(out$table$PValue)
library(rtracklayer)
export(con='LNA_h3k4me3.bed', db.regions[out$table$FDR<=0.05])
```

## In both

We repeat this process for the LNA/RNAi combined results:

```{r}
out <- consolidateSizes(all.regions, all.combined, 
    merge.args=list(max.width=10e3, tol=100))
```

We examine the number of regions that change, and in which direction.

```{r}
table(out$table$FDR <= 0.05, out$table$direction)
```

We save this result to file.

```{r}
best <- getBestTest(unlist(out$id), do.call(rbind, all.combined))
output <- data.frame(as.data.frame(out$region)[,1:3],
    out$table, Best.logFC=best$logFC, surrounds)
output <- output[order(output$PValue),]    
write.table(file="Both_h3k4me3_regions.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
```

Also the significant hits to a BED file for viewing.

```{r}
db.regions <- out$region
db.regions$score <- -10*log10(out$table$PValue)
library(rtracklayer)
export(con='Both_h3k4me3.bed', db.regions[out$table$FDR<=0.05])
```

# Reporting DB genes

## Defining the regions

We obtain genic regions from the Ensembl annotation:

```{r}
hs.genes <- genes(EnsDb.Hsapiens.v86)
length(hs.genes)
```

Fixing the sequence levels:

```{r}
seqlevels(hs.genes) <- paste0("chr", seqlevels(hs.genes))
seqlevels(hs.genes)
```

Getting the gene symbols.

```{r}
anno <- data.frame(Ensembl=hs.genes$gene_id, Symbol=hs.genes$gene_name)
head(anno)   
```

We also create a function to consolidate the best scores across sizes.

```{r}
allBest <- function(ids, results, FUN=getBestOverlaps) {
    out <- mapply(FUN=FUN, ids, results, SIMPLIFY=FALSE)
    pvals <- lapply(out, "[[", "PValue")
    lfcs <- lapply(out, "[[", "logFC")
    all.p <- do.call(cbind, pvals)
    all.p[is.na(all.p)] <- 1
    i <- max.col(-all.p)
    unname(do.call(cbind, lfcs)[cbind(seq_along(i), i)])
}
```

## RNAi

Consolidating results across window sizes:

```{r}
suppressWarnings({
    out <- consolidateSizes(all.regions, all.rnai, region=hs.genes)
})
```

We examine the number of regions that change, and in which direction.

```{r}
table(out$table$FDR <= 0.05, out$table$direction)
```

We save this result to file.

```{r}
output <- data.frame(anno, out$table, best.logFC=allBest(out$id, all.rnai))
output <- output[order(output$PValue),]    
output <- output[!is.na(output$PValue),]
write.table(file="RNAi_h3k4me3_genes.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)
```

## LNA

Consolidating results across window sizes:

```{r}
suppressWarnings({
    out <- consolidateSizes(all.regions, all.lna, region=hs.genes)
})
```

We examine the number of regions that change, and in which direction.

```{r}
table(out$table$FDR <= 0.05, out$table$direction)
```

We save this result to file.

```{r}
output <- data.frame(anno, out$table, best.logFC=allBest(out$id, all.lna))
output <- output[order(output$PValue),]    
output <- output[!is.na(output$PValue),]
write.table(file="LNA_h3k4me3_genes.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
```

## Both

Consolidating results across window sizes:

```{r}
suppressWarnings({
    out <- consolidateSizes(all.regions, all.combined, region=hs.genes)
})
```

We examine the number of regions that change, and in which direction.

```{r}
table(out$table$FDR <= 0.05, out$table$direction)
```

We save this result to file.

```{r}
output <- data.frame(anno, out$table, best.logFC=allBest(out$id, all.combined))
output <- output[order(output$PValue),]    
output <- output[!is.na(output$PValue),]
write.table(file="Both_h3k4me3_genes.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
```

# Session information

```{r}
sessionInfo()
```
