---
title: Consolidating DB analyses for #_MARK_# cut-and-run 
author: Aaron Lun
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    fig_caption: yes
---

```{r, echo=FALSE, message=FALSE, results="hide"}
library(BiocStyle)
knitr::opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
```

# Reading in the results

Here we consolidate the various window-based analyses into a single set of results.
First, we define the files to load in:

```{r}
library(csaw)
all.files <- list.files(full=TRUE, pattern="^#_MARK_#_.*rds$")
all.files
```

And we load in the results:

```{r}
all.results <- lapply(all.files, readRDS)
all.regions <- lapply(all.results, "[[", i="windows")
all.rnai <- lapply(all.results, "[[", i="RNAi")
all.lna <- lapply(all.results, "[[", i="LNA")
```

We apply an IUT on the one-sided p-values.
This involves converting each p-value into two one-sided values, performing IUTs on each side, and then effectively performing a Bonferoni correction to combine them again.

```{r}
all.combined <- vector("list", length(all.results))
for (idx in seq_along(all.rnai)) {
    cur.rnai <- all.rnai[[idx]]
    cur.lna <- all.lna[[idx]]

    up.rnai <- cur.rnai$logFC > 0
    left.rnai <- ifelse(up.rnai, cur.rnai$PValue/2, 1-cur.rnai$PValue/2)
    right.rnai <- ifelse(!up.rnai, cur.rnai$PValue/2, 1-cur.rnai$PValue/2)

    up.lna <- cur.lna$logFC > 0
    left.lna <- ifelse(up.lna, cur.lna$PValue/2, 1-cur.lna$PValue/2)
    right.lna <- ifelse(!up.lna, cur.lna$PValue/2, 1-cur.lna$PValue/2)

    left.p <- pmax(left.rnai, left.lna)
    right.p <- pmax(right.rnai, right.lna)
    all.combined[[idx]] <- data.frame(logFC=(cur.rnai$logFC + cur.lna$logFC)/2,
        PValue=pmin(left.p, right.p, 0.5)*2)
}
```

# Reporting DB regions

## Consolidating regions

We consolidate all of these results into a single set of DB regions, with a maximum region width of 10 kbp.

```{r}
merged <- consolidateWindows(all.regions, 
    merge.args=list(max.width=10e3, tol=100))
merged$region
```

## RNAi

We compute the consolidated statistics for the RNAi comparison.

```{r}
tabcom <- consolidateTests(merged$id, all.rnai, merged$weight)
head(tabcom)
```

We examine the number of regions that change, and in which direction.

```{r}
table(tabcom$FDR <= 0.05, tabcom$direction)
```

We also get the best log-fold changes in each region.

```{r}
best <- consolidateTests(merged$id, all.rnai, merged$weight,
    FUN=getBestTest)
head(best)
```

Adding annotation.

```{r}
library(org.Hs.eg.db)
library(EnsDb.Hsapiens.v86)
fiddling <- merged$region
seqlevels(fiddling) <- paste0("tmp_", seqlevels(fiddling)) # avoid clash with yeast X.
seqlevels(fiddling) <- sub("^tmp_chr", "", seqlevels(fiddling))
suppressWarnings({
    surrounds <- detailRanges(orgdb=org.Hs.eg.db, txdb=EnsDb.Hsapiens.v86, 
        fiddling, key.field="ENSEMBL")
})
```

We save this result to file.

```{r}
output <- data.frame(as.data.frame(merged$region)[,1:3],
    tabcom, Best.logFC=best$logFC, surrounds)
output <- output[order(output$PValue),]    
write.table(file="RNAi_#_MARK_#_regions.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
head(output)
```

Also the significant hits to a BED file for viewing.

```{r}
db.regions <- merged$region
db.regions$score <- -10*log10(tabcom$PValue)
library(rtracklayer)
export(con='RNAi_#_MARK_#.bed', db.regions[tabcom$FDR<=0.05])
```

## LNA

We repeat this process for LNA:

```{r}
tabcom <- consolidateTests(merged$id, all.lna, merged$weight)
head(tabcom)
```

We examine the number of regions that change, and in which direction.

```{r}
table(tabcom$FDR <= 0.05, tabcom$direction)
```

We also get the best log-fold changes in each region.

```{r}
best <- consolidateTests(merged$id, all.lna, merged$weight,
    FUN=getBestTest)
head(best)
```

We save this result to file (no need to recompute annotation, as the regions are the same).

```{r}
output <- data.frame(as.data.frame(merged$region)[,1:3],
    tabcom, Best.logFC=best$logFC, surrounds)
output <- output[order(output$PValue),]    
write.table(file="LNA_#_MARK_#_regions.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
```

Also the significant hits to a BED file for viewing.

```{r}
db.regions <- merged$region
db.regions$score <- -10*log10(tabcom$PValue)
library(rtracklayer)
export(con='LNA_#_MARK_#.bed', db.regions[tabcom$FDR<=0.05])
```

## In both

We repeat this process for the LNA/RNAi combined results:

```{r}
tabcom <- consolidateTests(merged$id, all.combined, merged$weight)
head(tabcom)
```

We examine the number of regions that change, and in which direction.

```{r}
table(tabcom$FDR <= 0.05, tabcom$direction)
```

We also get the best log-fold changes in each region.

```{r}
best <- consolidateTests(merged$id, all.lna, merged$weight,
    FUN=getBestTest)
head(best)
```

We save this result to file.

```{r}
output <- data.frame(as.data.frame(merged$region)[,1:3],
    tabcom, Best.logFC=best$logFC, surrounds)
output <- output[order(output$PValue),]    
write.table(file="Both_#_MARK_#_regions.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
```

Also the significant hits to a BED file for viewing.

```{r}
db.regions <- merged$region
db.regions$score <- -10*log10(tabcom$PValue)
library(rtracklayer)
export(con='Both_#_MARK_#.bed', db.regions[tabcom$FDR<=0.05])
```

# Reporting DB genes

## Defining the regions

We obtain genic regions from the Ensembl annotation:

```{r}
hs.genes <- genes(EnsDb.Hsapiens.v86)
hs.genes[,2]
```

We also kick out each gene by 3 kbp at the 5' end, to capture binding at the putative promoeter region.

```{r}
suppressWarnings({
    hs.genes <- resize(hs.genes, width(hs.genes)+3e3, fix="end")
})
hs.genes[,2]
```

Fixing the sequence levels:

```{r}
seqlevels(hs.genes) <- paste0("chr", seqlevels(hs.genes))
head(seqlevels(hs.genes))
```

Getting the gene symbols.

```{r}
anno <- data.frame(Ensembl=hs.genes$gene_id, Symbol=hs.genes$gene_name)
head(anno)   
```

Finding the windows that overlap each gene:

```{r}
suppressWarnings({
    molaps <- consolidateWindows(all.regions, region=hs.genes)
})
```

## RNAi

We compute combined statistics for each gene:

```{r}
tabcom <- consolidateOverlaps(molaps$olap, all.rnai, molaps$weight)
head(tabcom)    
```

We examine the number of regions that change, and in which direction.

```{r}
table(tabcom$FDR <= 0.05, tabcom$direction)
```

We also compute the best statistics for each gene.

```{r}
tabbest <- consolidateOverlaps(molaps$olap, all.rnai,
    molaps$weight, FUN=getBestOverlaps)
head(tabbest)
```

We save this result to file.

```{r}
output <- data.frame(anno, tabcom, best.logFC=tabbest$logFC)
output <- output[order(output$PValue),]    
output <- output[!is.na(output$PValue),]
write.table(file="RNAi_#_MARK_#_genes.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)
```

## LNA

Consolidating results across window sizes:

```{r}
tabcom <- consolidateOverlaps(molaps$olap, all.lna, molaps$weight)
head(tabcom)    
```

We examine the number of genes that change, and in which direction.

```{r}
table(tabcom$FDR <= 0.05, tabcom$direction)
```

We also compute the best statistics for each gene.

```{r}
tabbest <- consolidateOverlaps(molaps$olap, all.lna,
    molaps$weight, FUN=getBestOverlaps)
head(tabbest)
```

We save this result to file.

```{r}
output <- data.frame(anno, tabcom, best.logFC=tabbest$logFC)
output <- output[order(output$PValue),]    
output <- output[!is.na(output$PValue),]
write.table(file="LNA_#_MARK_#_genes.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
```

## Both

Consolidating results across window sizes:

```{r}
tabcom <- consolidateOverlaps(molaps$olap, all.combined, molaps$weight)
head(tabcom)    
```

We examine the number of genes that change, and in which direction.

```{r}
table(tabcom$FDR <= 0.05, tabcom$direction)
```

We also compute the best statistics for each gene.

```{r}
tabbest <- consolidateOverlaps(molaps$olap, all.combined,
    molaps$weight, FUN=getBestOverlaps)
head(tabbest)
```

We save this result to file.

```{r}
output <- data.frame(anno, tabcom, best.logFC=tabbest$logFC)
output <- output[order(output$PValue),]    
output <- output[!is.na(output$PValue),]
write.table(file="Both_#_MARK_#_genes.tsv", output,
    row.names=FALSE, sep="\t", quote=FALSE)        
```

# Session information

```{r}
sessionInfo()
```
